Seismic velocity estimation and the Canadian crust

### 2014-04-14 paper/bpostlethwaite_Thesis.tex
thunk

Our knowledge of the continental crust in Canada, in particular its thickness and structure in depth, has advanced greatly through Lithoprobe, a now complete national geoscience project spearheaded by active-source seismic profiling (Clowes et al., 1999). The full suite of Lithoprobe seismic profiles form a near-continuous transect from the Pacific to the Atlantic Oceans, and has enabled many insights into crustal structure and genesis. Recent summary studies utilizing Lithoprobe data include the analysis of {\it Eaton} (2005) which considers seismic observations in light of several hypotheses for the origin of the continental Moho: relic Moho, underplating, metamorphic front and regional d\'{e}collement. Seismic evidence of preserved island arc Moho in Cascadia supports the first hypothesis whereas the presence of high-velocity, lower-crustal xenoliths in southern Alberta favours thick mafic underplating. Eclogite mineral assemblages forming a deep crustal root below the Grenville Front lends support to the third hypothesis, that at least in some places, continental Moho forms by large-scale eclogitization of the lower crust. Finally, seismic imaging of subhorizontal reflections separated by dipping reflection fabrics beneath the southern Canadian Cordillera suggests decoupling of the upper crust from the mantle during deformation and is taken as evidence for the fourth hypothesis, that an overprinted Moho may serve as a subhorizontal regional d\'ecollement. A subsequent study by {\it Cook et al.} (2009) argues that the continental Moho has varied and complex origins and that its imprint evolves through the effects of temperature, pressure, deformation and fluid migration.

>>>>

Our knowledge of the continental crust in Canada, in particular its thickness and structure in depth, has advanced greatly through Lithoprobe, a now complete national geoscience project spearheaded by active-source seismic profiling (Clowes et al., 1999). The full suite of Lithoprobe seismic profiles form a near-continuous transect from the Pacific to the Atlantic Ocean, and has enabled many insights into crustal structure and genesis. Recent summary studies utilizing Lithoprobe data include the analysis of {\it Eaton} (2005) which considers seismic observations in light of several hypotheses for the origin of the continental Moho: relic Moho, underplating, metamorphic front and regional d\'{e}collement. Seismic evidence of preserved island arc Moho in Cascadia supports the first hypothesis whereas the presence of high-velocity, lower-crustal xenoliths in southern Alberta favours thick mafic underplating. Eclogite mineral assemblages forming a deep crustal root below the Grenville Front lends support to the third hypothesis, that at least in some places, continental Moho forms by large-scale eclogitization of the lower crust. Finally, seismic imaging of subhorizontal reflections separated by dipping reflection fabrics beneath the southern Canadian Cordillera suggests decoupling of the upper crust from the mantle during deformation and is taken as evidence for the fourth hypothesis, that an overprinted Moho may serve as a subhorizontal regional d\'ecollement. A subsequent study by {\it Cook et al.} (2009) argues that the continental Moho has varied and complex origins and that its imprint evolves through the effects of temperature, pressure, deformation and fluid migration.
